\documentclass[11pt,draft,phd]{osudiss-2}

% cli to extract pages from pdf
% qpdf input.pdf --pages . 1-10 -- output.pdf

\usepackage[english]{babel}
\usepackage{microtype}
\usepackage{environ}
\usepackage{amsmath}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{doi}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{scrextend} % for \footref
\usepackage{hyperref}
\usepackage[nameinlink,capitalize]{cleveref}
\usepackage{siunitx}
\usepackage{enumitem}
\usepackage{floatrow}
\usepackage{bm}
\usepackage{mathtools}

%\usepackage[acronym, section=chapter]{glossaries}
%\makeglossaries
%\include{acronyms}

\usepackage{lipsum}

% style stuff
\bibliographystyle{apsrev4-1-custom}
\addto{\captionsenglish}{\renewcommand{\bibname}{References}}
\hypersetup{hidelinks}

% custom commands
\DeclareMathOperator{\Tanh}{\mathbf{tanh}}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}

% environment for figure re-use
% https://tex.stackexchange.com/a/225075
% \begin{reusefigure}[<float spec>]{<ref>}
%\newenvironment{reusefigure}[2][tbp]
%  {\addtocounter{figure}{-1}%
%   \renewcommand{\theHfigure}{dupe-fig}% If you're using hyperref
%   \renewcommand{\thefigure}{\ref{#2}}% Figure counter is \ref
%   \renewcommand{\addcontentsline}[3]{}% Avoid placing figure in LoF
%   \begin{figure}[#1]}
%  {\end{figure}}

% use dcolumn
\newcolumntype{d}{D{.}{.}{-1}}

% full references with names
\newcommand*{\fullref}[1]{\hyperref[{#1}]{\Cref*{#1} \nameref*{#1}}}

\title{Essential Reservoir Computing}
\author{Aaron Griffith}
\advisorname{Daniel J. Gauthier}
\degree{Doctor of Philosophy}
\member{Amy Connolly}
\member{Ciriyam Jayaprakash}
\member{Gregory Lafyatis}
\authordegrees{B.S.}
\graduationyear{2021}
\unit{Graduate Program in Physics}

\begin{document}

\frontmatter

\begin{abstract}
  % abstract: 500 words or less

  \emph{Reservoir computing} (RC) is a machine learning method
  especially well suited to solving physical problems, by using an
  internal dynamic system known as a 'reservoir'. Many systems are
  suitable for use as an internal reservoir. A common choice is an
  \emph{echo state network} (ESN), a network with recurrent
  connections that gives the RC a memory which it uses to efficiently
  solve many time-domain problems such as forecasting chaotic systems
  and hidden state inference. However, constructing an ESN involves a
  large number of poorly-understood meta-parameters, and the properties
  that an ESN must have to solve these tasks well are largely unknown.

  In this dissertation, I explore what parts of an RC are absolutely
  necessary. I build ESNs that perform well at system forecasting
  despite an extremely simple internal network structure, without any
  recurrent connections at all, breaking one of the most common rules
  of ESN design. These simple reservoirs indicate that the role of the
  reservoir in the RC is only to remember a finite number of
  time-delays of the RCs input, and while a complicated network can
  achieve this, in many cases a simple one achieves this as well.

  I then build upon a recent proof of the equivalence between a
  specific ESN construction and the nonlinear vector auto-regression
  (NVAR) method with my collaborators. The NVAR is an RC boiled down
  to its most essential components, taking the necessary time-delay
  taps directly rather than relying on an internal dynamic
  reservoir. I demonstrate these RCs-without-reservoirs on a variety
  of classical RC problems, showing that in many cases an NVAR will
  perform as well or better than an RC despite the much simpler
  method. I then conclude with an example problem that highlights a
  remaining unsolved issue in the application of NVARs, and then look
  to a possible future where NVARs may supplant RCs.
\end{abstract}

\dedication{For my brother Nathan, and my parents Gregory and Mary Lea.}

\begin{acknowledgments}
  This work would not have been possible without the committed support
  of my advisor, Daniel~J.~Gauthier, or my insightful collaborators,
  including Wendson~A.~S.~Barbosa, Erik~Bollt, Daniel~Canaday, and Andrew~Pomerance.

  In addition, many discussions over the years have
  re-contextualized my understanding of reservoir computers. I would
  specifically like to thank Michelle~Girvan, Brian~Hunt, and Edward~Ott.
\end{acknowledgments}

\begin{vita}
  \dateitem{2009-2014}{B.S., Mathematics and Physics \\ The Ohio State University}
  \dateitem{2015-2016}{Graduate Teaching Associate \\ Department of Physics \\ The Ohio State University}
  \dateitem{2016-2021}{Graduate Research Associate \\ Department of Physics \\ The Ohio State University}

  \begin{publist}
    \pubitem{D.~Canaday, A.~Griffith, and D.~J.~Gauthier, ``Rapid time series prediction with a hardware-based reservoir computer,'' \href{https://doi.org/10.1063/1.5048199}{Chaos: An Interdisciplinary Journal of Nonlinear Science \textbf{28}, 123119 (2018)}.}
    \pubitem{A.~Griffith, A.~Pomerance, and D.~J.~Gauthier, ``Forecasting chaotic systems with very low connectivity reservoir computers,'' \href{https://doi.org/10.1063/1.5120710}{Chaos: An Interdisciplinary Journal of Nonlinear Science \textbf{29}, 123108 (2019)}.}
    %\pubitem{W.~A.~S.~Barbosa, A.~Griffith, G.~E.~Rowlands, L.~C.~G.~Govia, G.~J.~Ribeill, M.-H.~Nguyen, T.~A.~Ohki, and D.~J.~Gauthier, ``Symmetry-aware reservoir computing,'' \href{https://arxiv.org/abs/2102.00310}{(2021), \\ arXiv:2102.00310 [cs.NE]}.}
    %\pubitem{D.~J.~Gauthier, E.~Bollt, A.~Griffith, W.~A.~S.~Barbosa, ``Next generation reservoir computing,'' \href{https://arxiv.org/abs/2106.07688}{(2021), arXiv:2106.07688 [cs.LG]}.}
  \end{publist}

  \begin{fieldsstudy}
    \majorfield{Physics}
  \end{fieldsstudy}
\end{vita}

\tableofcontents 

\clearpage
\listoffigures 

\clearpage
\listoftables 

%\clearpage
%\PrintListofAbbreviations{List of Abbreviations}

\mainmatter
\include{1-introduction}
\include{2-reservoir-computing}
\include{3-low-connectivity}
\include{4-nvar}
\include{5-nvar-application}
\include{6-conclusion}

\backmatter
%\nocite{*}
\bibliography{thesis}

\appendix
\include{A1-systems}

\end{document}
